Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-09-17 writeups/dissertation/dissertation.tex
foo

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The task of choosing a hashtag for a tweet and tagging a post on StackOverflow was framed as a declarative memory retrieval problem.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random permutation vector-based model.
  The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, the common natural language processing method to attenuate low-predictor words in the dataset was linked to ACT-R's attentional weight term,
  and an entropy-weighting technique was developed that improved model accuracy by more efficiently allocating limited attentional resources to the predictive words in the posts and tweets.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order,
  which shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

\section{Motivation}

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human declarative memory.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

@@

But this approach is problematic for two main reasons:
First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.

@@

The work for this proposal will expand on previous work that has tested retrieval models on large-scale datasets.
It will focus specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and random permutations) can be improved, by testing these models on two real-world tagging tasks.
One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
Large-scale datasets will be used throughout the model development and evaluation process.
This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.

\subsection{Motivation for Task}

@@

So how can we ensure that users are connected to the information streams that they care most about?
One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
If we have a better understanding of a user's goals, then we can tailor the information streams provided to the ones that are most in line with these goals.
For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out.
Researchers are currently looking for ways to automate the goal-identification task for more general search queries like those generated for Google \parencites{Jansen2008}{Lee2005}.

@@

These hashtags relate to a user's goals and interests because they are a form of human-based document tagging \parencite{Chang2010}.
Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
Twitter users, for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
With that understanding, we can tailor the system to suggest hashtags that they may not yet know about but are of likely interest to them.
We can help aid in their discovery of fresh and relevant information that matches their goals on the site.

@@

The core research question motivating this proposal is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
This process can be framed as a memory retrieval problem.
That is, each user on a social media site has created content that contain a set of tags within the context of each post.
The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
At least two of the current memory retrieval models are based on a co-occurrence methodology, and I will be comparing these two models for hashtag prediction:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

@@

An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
However, it is unclear and unexplored how word order can be represented in the ACT-R declarative memory theory.

ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
This decay rate equation formalizes how recency and frequency of hashtag use relate to the prior probability that a hashtag will be retrieved.
However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.

@@

\section{Prior Research}

@@

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the brain (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up long-term declarative memory retrieval.
The equations that make up this system can be derived through a rational analysis of long-term memory retrieval.
That is, given the task of retrieving a chunk of information from long-term declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

@@

\begin{table}[!ht]
  \caption{ACT-R declarative memory model}
  \label{tabACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Constant Base Level &	$B_{i} = log \frac{p(i)}{p(\overline{i})}$ \\
      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)}$ \\
      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
      \hline
    \end{tabu}
  }
\end{table}

@@

Base-level activation reflects the log prior odds of needing an observed chunk again.
The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its current frequency and recency of use.

@@

Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
Context for the Twitter domain are the words in a tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption has been both mentioned and used when deriving the $S_{ji}$ equation \parencite{Anderson1989}, as well as in recent research working with large-scale datasets \parencite{Stanley2013}.

@@

Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
The PMI index is included in Equation \eqref{eqPMI}.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is a technique that has mathematical roots in factor analysis and also measures strength of association between two terms.
The process starts with a set of N documents.
A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

Usually information is lost when the data are restructured into a smaller set of latent dimensions.
However, it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
Model results were on par with a large sample of applicants to american universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
Afterwards further reductions in the dimensionality of the data began to degrade performance.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, Stackoverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., chunk) as a vector.
The representation of the concept is distributed across all elements of the vector.
This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.

@@

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.
However, decoding does require an additional decoding step that resonance does not require, and depending on the decoding operation, this step may be computationally expensive.

@@

\textcite{Jones2007} created the BEAGLE vector-based memory representation.
Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
This model uses circular convolution and circular correlation as a way to encode and decode word order.
BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (\num{55.6}\% compared to \num{55.3}\%).
Also, BEAGLE with unordered context and word order was more correlated to the WordNet synonym database than LSA (\num{-.311} compared to \num{-.165}).
Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.

@@

A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.

@@

With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.

@@

@@

Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the benefits of one over the other.
For example, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.
More recent tweets on a news site like Twitter should be more relevant, for example.
\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.

ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
This type of information on a user's prior behavior is not being utilized with vector-based models currently.

@@

One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
It may be that multiple co-occurrence matrices must be used to represent ordered and unordered co-occurrences separately.
However, it would be more parsimonious if these matrices could be combined into a single matrix representation, much like how this information is combined for vector-based models.

@@

@@

A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
These general recommendation systems have grown in need with the growth of the web.
Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.

@@

Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.

@@

The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not publicized.
StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags, as the algorithm has not yet been deployed for the busiest StackOverflow site.
The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.

>>>>

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

\section{Introduction}

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

@@

However, this approach is problematic for two main reasons:
First, it means that the declarative memory system for every developed model must be hand crafted and customized for that model, which takes enormous time and effort.
Second, it means that modelers are continuing to use small sample sizes for the declarative memory components in their models.

@@

The work in this study expands on previous work that has tested retrieval models on large-scale datasets.
It focuses specifically on how two state-of-the-art declarative memory retrieval models (ACT-R Bayesian and vector-based) can be improved by testing these models on two real-world tagging tasks.
One primary objective is to compare the random permutation vector-based model to ACT-R's declarative memory theory when large datasets are used for each model. 
Another objective is to improve each model, and measure how incorporating the strengths of each model into the other (i.e., modifying model components) influences accuracy.
For example, it is currently unexplored how a vector-based model can be modified to retrieve results that are based not only on context, but also on the prior odds that a particular item in memory will be needed again.
Large-scale datasets were used throughout the model development and evaluation process.
This provides an ideal testing environment for constraints and assumptions for each theory to be rapidly tested and new model components to be hypothesized and quickly evaluated.

\subsection{The Task: Predicting User-Generated Tags}

@@

So how can we ensure that users are connected to the information streams that they care most about?
One promising approach is to model and predict the user's goals when using and searching these social media sites \parencite{Rose2004}.
If we have a better understanding of a user's goals, then we can tailor the information streams to the content that is most in line with these goals.
For example, Twitter might suggest other users and hashtags for the user to follow, Facebook could suggest pages to like, Yelp could recommend businesses to check out.
Researchers are currently looking for ways to automate the goal-identification task for more general search queries like those generated for Google \parencites{Jansen2008}{Lee2005}.

@@

These hashtags relate to a user's goals and interests because they are a form of human-based document tagging \parencite{Chang2010}.
Further, people are using hashtags as a way to create information streams on social media sites \parencite{Kwak2010}.
Twitter users, for example, commonly create and use hashtags for upcoming political events that interest them, such as debates and races \parencite{Diakopoulos2010}.
So if we can predict what hashtags a user is interested in, we have a better understanding of the information streams that they care about.
We can use that information to correctly suggest hashtags that they have previously used when they are generating new content.

Further, if a particular information stream has multiple hashtags associated with it (common for Twitter), these semantically similar hashtags can also be recommended to users.
So we can tailor the system to suggest hashtags that the user may not yet know about but are of likely interest to them and growing in popularity across the site.
We can help aid in their discovery of fresh and relevant information that matches their goals on the site.

@@

The core research question motivating this proposal is:
What kinds of cognitively-plausible models can best predict the tags that a user on a social media site uses?
This question can be tested by generating a prediction for the most likely used hashtag whenever a user is about to generate a hashtag, and then comparing the user's chosen hashtag to the model's prediction.
This process can be framed as a memory retrieval problem.
The process of suggesting a relevant hashtag to a user can be thought of as a memory retrieval request for a hashtag, given prior hashtag use and current context.
That is, each user on a social media site has created content that contains a set of tags within the context of each post.

Co-occurrence-based modeling has been shown as a potentially useful approach when predicting Twitter hashtag use \parencite{Efron2010}.
Several memory retrieval models are based on this broad methodology, where a count is maintained of the times each contextual element (such as a word in a post) co-occur with a tag.
At least two of the current memory retrieval models are based on a co-occurrence methodology:
ACT-R's declarative memory (DM) retrieval theory and random permutation vector-based memory systems.

@@

An advantage of vector-based models is that word order can be easily incorporated into the single co-occurrence representation \parencite{Jones2007}.
Word order information on sites like Twitter may contain highly predictive pieces of information, as it is likely that specific words immediately precede specific hashtags.
Word order has been incorporated into ACT-R sentence-parsing models previously, but this has not been done using declarative memory.
Rather, production rules were defined that ensured that the model followed a set of rules (e.g., left-to-right reading) when parsing sentences \parencite{Lewis2006}.
It is certainly true that at least some production rules are needed when choosing words and creating content. 
However, it is unexplored if some of the process of sentence construction and word selection can be represented in the declarative memory system.

ACT-R's declarative memory system places strong constraints on how a user's prior knowledge and experience influences the likelihood that a particular memory item is retrieved \parencite{Anderson2007}.
This decay rate equation formalizes how recency and frequency of memory use relate to the prior probability that a memory will be retrieved.
However, it is unexplored how a user's prior knowledge should influence retrieval for vector-based models.

@@

\section{Theoretical Background}

@@

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
The equations that make up this system were derived from a rational analysis of long-term memory retrieval.
That is, given the task of retrieving a chunk of information from long-term declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

@@

\begin{table}[!ht]
  \caption{ACT-R declarative memory model}
  \label{tabACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Constant Base Level &	$B_{i} = log \frac{p(i)}{p(\overline{i})}$ \\
      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
      \hline
    \end{tabu}
  }
\end{table}

@@

Base-level activation reflects the log prior odds of needing an observed chunk again.
The default way to calculate these log prior odds is to use the standard base-level equation in Table \ref{tabACTRModel}.
This equation formalizes how log prior odds are a function of both frequency and recency of prior exposure to a particular chunk.
Chunks used more frequently (either through exposure or from a retrieval) are more likely to be needed for retrieval again.
However, as time progresses and a particular chunk is no longer used, the activation for that chunk decays.
In this way the standard base-level equation formalizes the time dynamics of the retrieval system, where a chunk's base-level activation evolves over time, depending on its frequency and recency of use.

@@

Strength of association ($S_{ji}$) reflects the amount of log-odds adjustment to the activation of a chunk, given the current context (i.e., external environment and internal state).
Context for the StackOverflow domain for example is represented as each word in the title and body of a post.
Context for the Twitter domain are the words in a tweet.
Association strength between a chunk in memory and a single contextual element can be computed directly by calculating its context-adjusted odds ratio:
The likelihood a chunk occurred with the current context ($p(i|j)$) over the likelihood that the chunk occurred in any of the other contexts ($p(i|\overline{j})$).
For large datasets, the likelihood that a chunk occurs in any particular context reaches near-zero values ($p(j) \Rightarrow 0$).
So it can be assumed that the likelihood that a chunk occurs in any context but one ($p(i|\overline{j})$) is equivalent to the likelihood that a chunk occurs in any context ($p(i)$).
This assumption was described when deriving the $S_{ji}$ equation \parencite{Anderson1989}, and used in recent research working with large-scale datasets \parencites{Stanley2013,Farahat2004,Douglass2010}.

@@

Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
The PMI index is presented as Equation \eqref{eqPMI}.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
The process starts with a set of N documents.
A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

Usually information is lost when the data are restructured into a smaller set of latent dimensions.
However, it was shown that this process actually improves prediction and reduces noise by allowing the frequency count of highly similar concepts (words) to be pooled together and represented by a single dimension.
\textcite{Landauer1997} tested the LSA model on the Test of English as a Foreign Language (TOEFL) task of choosing the best of four possible synonyms to a target word.
Model results were on par with a large sample of applicants to American universities from non-English speaking countries (\num{64.4}\% compared to \num{64.5}\%).
Model performance also improved when the original dimensionality of the data was reduced, but only up to a point.
Afterwards further reductions in the dimensionality of the data began to degrade performance.

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is aimed to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

@@

``Holographic'' memory systems \parencite{Plate1995} or vector-based memory systems represent a concept (i.e., a chunk) as a vector.
The representation of the concept is distributed across all elements of the vector.
This is somewhat similar to taking a column (i.e., a tag) on a word co-occurrence matrix and viewing the distribution of co-occurrence counts across the rows in the column as the representation of that concept.
However, vector-based memory systems represent information in a much more compact way than a full word co-occurrence matrix.

@@

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.

@@

\textcite{Jones2007} created the BEAGLE vector-based memory representation.
Each memory vector ($m_{i}$) is a summed representation of unordered and ordered word co-occurrences ($c_{i} + o_{i}$).
This model uses circular convolution and circular correlation as a way to encode and decode word order.
BEAGLE performed similar to LSA when trained on the TASA and tested on the TOEFL (\num{55.6}\% compared to \num{55.3}\%).
Further, it was shown that incorporating word order information into the same representation resulted in minimal data loss (less than 1 percent of total predictive variance).
This model provided an efficient algorithm for using a single representation for unordered and ordered information that showed performance similar to if not better than LSA.

@@

A formal description of the random permutation model in \textcite{Sahlgren2008} is included in Table \ref{tabRandPermModel}.
With random permutations, each word environment vector ($e_{i}$) is a large sparse vector of zeros with a few one and negative one values in random locations. 
In order to create a new environment vector for a word that preceded another word in a sentence, that word's environment vector is shifted to the left one position.
This produces a new environment vector ($e_{i^{-1}}$) for the combination of the word and the position that is uncorrelated with the original word's environment vector and all other environment vectors.

@@

@@

The fan effect describes a behavior of human memory where memory items are more difficult to retrieve if they have occurred in a larger number of different contexts.
\textcite{Rutledge2008} used \textcite{Anderson1974} classic experimental paradigm to test the fan effect.
In the task, participants are exposed to a set of person and place pairs (e.g., ``the hippie is in the park'').
They are later asked to recall if a particular person and place pair was presented to them before.
The main manipulation is the ``fan'' of the person and place for each pair.
The fan is defined as the number of different contexts that each item in the pair is presented in.
For example, if items were presented to the participant where the ``hippie'' was both in the ``park'' and in the ``store'', then ``hippie'' would have a fan of two.
Recall accuracy decrease and time to recall increase as total fan of the person and pair increases.

Both the ACT-R declarative memory theory and vector-based models produce the fan effect due to the effect that fan has on the co-occurrence matrix.
As a particular word (e.g., ``hippie'') is seen in more and more contexts, the co-occurrence count of the word across all contexts is spread across multiple contexts.
This means that when a retrieval for a context given the word occurs, multiple contexts have relatively high activation.
This makes it more difficult to test if a signal (e.g., if a word appears in a specific context) is higher than the background noise (e.g., all other contexts that the word has appeared in),
which makes the retrieval more uncertain.

Both vector-based and Bayesian models produce the fan effect since they both are based on a co-occurrence matrix.
However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.

@@

Vector-based models and ACT-R's current Bayesian DM model are not identical however, and these differences need to be addressed in order to better evaluate the relative benefits of the two models. 
In particular, the frequency and recency of retrievals for each chunk are not currently used for vector-based models when calculating activation.

The frequency and recency of how content has been used in the past can be a strong predictor of the future importance of that information. 
The recent content in tweets on a news site like Twitter should be more relevant, for example.
\textcite{Efron2011} tested this claim by evaluating a set of query likelihood models on returning relevant Twitter tweets from search queries.
They showed that a model incorporating recency information of the tweet produced more relevant results than models that did not pay attention to this information.

ACT-R's Bayesian model has a strong theory with respect to recency and frequency information.
It uses this information to calculate a chunk's prior activation (i.e., prior likelihood of being needed again, independent of current context).
Vector-based models compute activation based entirely on context, so chunks that resonate with context will be retrieved, irrespective to the frequency and recency of use of that chunk.
ACT-R's prior activation can also be computed for each user, so that hashtags that have been used frequently and recently for a particular user are rated with higher activation.
Vector-based models have focused on computing the associative strength of a chunk and specific contextual elements.
It is unexplored how accuracy changes for these models once the information about a user's prior behavior (e.g., ACT-R's base-level activation component) is added.

@@

One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
Although there are certainly ways to incorporate word order into the ACT-R DM model.
For example, an additional dimension could be added to the co-occurrence matrix representing word position.
But adding word-order information to ACT-R's default Bayesian model is not common.

@@

Using a correlation can be problematic due to range effects in observed co-occurrences.
If there are particular words for a tag that co-occur much more often than all of the other words with that tag, that single large value diminishes the power of all other words to contribute to the correlation calculation.
This does not happen with Bayesian retrieval models because each pair of word and tag values is normalized by the total number of words that appeared with that tag and total number of tags that appeared with that word
(see the strength of association equation in Table \ref{tabACTRModel}).
The high-frequency words are often removed from co-occurrence matrix for vector-based models to reduce this effect.

@@

A model that predicts a user-chosen hashtag is a specific type of a recommendation system.
General recommendation systems have grown in need with the growth of the web.
Users on a site need quick access to specific pieces of information, and the site contains much more information than they can process or care about.
Recommendation systems are used to tailor the information presented to what is most relevant to the user \parencite{Pazzani2007}.

@@

Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
The term frequency component measures the strength of association between a word and a document.
The inverse document frequency component measures how much information is contained within a specific word, independent of document.
ACT-R also uses two similar components to compute the activation of a word with a specific document.
The strength of association component of the ACT-R Bayesian co-occurrence model is a more sophisticated method for computing the term frequency component,
and ACT-R's attentional weight component of the declarative memory model can be used to measure the information content within each word.

@@

Recommendation models that attempt to predict a user's chosen hashtag have recently been created and tested for a few social media sites.
Google has even deployed a hashtag recommendation model to help users properly tag posts on their Google+ social media site \parencite{GoogleKeynote2013}.
A tag-suggestion tool has also been deployed for the \url{tex.stackexchange.org} and \url{superuser.com} StackExchange sites.
Models for Twitter and StackOverflow have been proposed and tested, but none have yet been deployed.
Recommendation models for StackOverflow, Google+, and Twitter will be described in more detail to follow.

@@

The exact tag-recommendation algorithm used for the Latex and SuperUser sites is not public.
After exploratory testing, StackExchange appears to be using something more sophisticated than simply looking for words in the title and body of the post that are tags.
The Latex and SuperUser sites are not nearly as popular as StackOverflow currently.
The sites have roughly \num{50000} and \num{189000} questions respectively compared to \num{5000000} questions on StackOverflow.
